---
title: "ovary"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages
```{r, include = FALSE}
library(tidyverse)
library(plyr)
library(reshape2)
library(lme4)
```

# import data from cleanData file
```{r}
# assumes working directory is where the clean1 datasheet (output from exp1clean.Rmd file) lives
getwd()
ovaryData <- read.csv("./clean1.csv", header = TRUE)

# make social treatment a factor with ordered levels
ovaryData$SocTreat <- factor(ovaryData$SocTreat, levels = c("solitary", "social")) 
```

# visualize data

## oocyte length
line graph of oocyte length
```{r}
# gather all oocyte lengths into a single column
plotData <- ovaryData %>%
  gather(Ooc11, Ooc12, Ooc13, Ooc14, Ooc21, Ooc22, Ooc23, Ooc24,
         key = "oocyte", 
         value = "oocLen")
# remove na values
plotData <- plotData[!is.na(plotData$oocLen),]

# count sample size
sampleSizes <- plotData %>% group_by(RelDayTreat, SocTreat) %>% tally()
sampleSizes

# summarize the data and calculate mean, sd, sem
summ <- ddply(plotData, 
                    c("RelDayTreat", "SocTreat"), 
                    summarise,
                    mean = mean(oocLen), 
                    sd = sd(oocLen),
                    sem = sd(oocLen)/sqrt(length(oocLen)))

# add rows with no data for social day 1 groups to make space in graph
addTimeGap <- data.frame(matrix(nrow = 1, ncol = ncol(summ)))
colnames(addTimeGap) <- colnames(summ)
addTimeGap$RelDayTreat <- "0 gas 9"
summ <- rbind(summ, addTimeGap)
# summ

# split summ into lines
natural <- summ[c(1, 2, 4, 6, 8, 10),]
natural$RelDayTreat <- as.character(natural$RelDayTreat)
natural <- natural[order(natural$RelDayTreat),]
natural$RelDayTreat <- as.factor(natural$RelDayTreat)
natural
premature <- summ[c(1, 3, 5, 11),]
premature$RelDayTreat <- as.character(premature$RelDayTreat)
premature <- premature[order(premature$RelDayTreat),]
premature$RelDayTreat <- as.factor(premature$RelDayTreat)
delayed <- summ[c(1, 2, 4, 6, 7, 9),]
delayed$RelDayTreat <- as.character(delayed$RelDayTreat)
delayed <- delayed[order(delayed$RelDayTreat),]
delayed$RelDayTreat <- as.factor(delayed$RelDayTreat)

#######

# line graph
ggplot() +
  geom_line(data = natural, aes(x = RelDayTreat, y = mean, group = 1), linetype = "dotted") +
  geom_point(data = natural, aes(x = RelDayTreat, y = mean)) +
  geom_line(data = premature, aes(x = RelDayTreat, y = mean, group = 1), linetype = "dotted") +
  geom_point(data = premature, aes(x = RelDayTreat, y = mean)) +
  geom_line(data = delayed, aes(x = RelDayTreat, y = mean, group = 1)) +
  geom_point(data = delayed, aes(x = RelDayTreat, y = mean)) +
  geom_errorbar(data = summ, aes(x = RelDayTreat, y = mean, fill = SocTreat, ymin = mean - sem, ymax = mean + sem),
                #position = position_dodge(.2),
                width = 0.1) +
  theme(text = element_text(size=15),
        #axis.text.x = element_text(angle = 90),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  scale_x_discrete(labels=c("1", "4\nearly", "7", "//", "1", "4\nlate", "7")) + 
  labs(x = "collection day at each nest stage", y = "oocyte length (mm)")
```

## resorbed oocytes
barplot of total number of resorbed oocytes
```{r}
# select only the data I'm interested in for this plot
#resorbPlotData <- select(ovaryData, RelDayTreat, SocTreat, nResOoc)
plotData <- ovaryData[!is.na(ovaryData$nResOoc),]

# count sample size
sampleSizes <- plotData %>% group_by(RelDayTreat, SocTreat) %>% tally()
sampleSizes

# summarize the data and calculate mean, sd, sem
summ <- ddply(plotData,
                    c("RelDayTreat", "SocTreat"),
                    summarise,
                    mean = mean(nResOoc),
                    sd = sd(nResOoc),
                    sem = sd(nResOoc)/sqrt(length(nResOoc)))

# add rows with no data for social day 1 groups to make space in graph
addTimeGap <- data.frame(matrix(nrow = 1, ncol = ncol(summ)))
colnames(addTimeGap) <- colnames(summ)
addTimeGap$RelDayTreat <- "0 gas 9"
summ <- rbind(summ, addTimeGap)
# summ

# split summ into lines
natural <- summ[c(1, 2, 4, 6, 8, 10),]
natural$RelDayTreat <- as.character(natural$RelDayTreat)
natural <- natural[order(natural$RelDayTreat),]
natural$RelDayTreat <- as.factor(natural$RelDayTreat)
natural
premature <- summ[c(1, 3, 5, 11),]
premature$RelDayTreat <- as.character(premature$RelDayTreat)
premature <- premature[order(premature$RelDayTreat),]
premature$RelDayTreat <- as.factor(premature$RelDayTreat)
delayed <- summ[c(1, 2, 4, 6, 7, 9),]
delayed$RelDayTreat <- as.character(delayed$RelDayTreat)
delayed <- delayed[order(delayed$RelDayTreat),]
delayed$RelDayTreat <- as.factor(delayed$RelDayTreat)

#######

# line graph
ggplot() +
  geom_line(data = natural, aes(x = RelDayTreat, y = mean, group = 1), linetype = "dotted") +
  geom_point(data = natural, aes(x = RelDayTreat, y = mean)) +
  geom_line(data = premature, aes(x = RelDayTreat, y = mean, group = 1), linetype = "dotted") +
  geom_point(data = premature, aes(x = RelDayTreat, y = mean)) +
  geom_line(data = delayed, aes(x = RelDayTreat, y = mean, group = 1)) +
  geom_point(data = delayed, aes(x = RelDayTreat, y = mean)) +
  geom_errorbar(data = summ, aes(x = RelDayTreat, y = mean, fill = SocTreat, ymin = mean - sem, ymax = mean + sem),
                #position = position_dodge(.2),
                width = 0.1) +
  theme(text = element_text(size=15),
        #axis.text.x = element_text(angle = 90),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  scale_x_discrete(labels=c("1", "4\nearly", "7", "//", "1", "4\nlate", "7")) + 
  labs(x = "collection day at each nest stage", y = "number of resorbed oocytes per queen")
```




########################################################################################################




# mixed models

### load packages
```{r, include = FALSE}
library(plyr)
library(ggplot2) 
library(ggpubr) 
library(lme4)
library(multcomp)
library(car)
library(AICcmodavg)
library(lmtest)
library(dplyr) 
library(tidyr)
library(cowplot)
library(MuMIn)
library(sjPlot)
library(effects)
```

## oocyte length

### restructure dataframe to use in model
```{r}
# gather all oocyte lengths into a single column
LenData <- ovaryData %>%
  gather(Ooc11, Ooc12, Ooc13, Ooc14, Ooc21, Ooc22, Ooc23, Ooc24,
         key = "oocyte", 
         value = "oocLen")
# remove na values
LenData <- LenData[!is.na(LenData$oocLen),]
LenData$DaysTreat <- as.factor(LenData$DaysTreat)
LenData
```

## decide on a distribution
normal is best
```{r}
# visualize distribution
ggdensity(LenData$oocLen)
hist(LenData$oocLen)

# normal distribution
qqp(LenData$oocLen, "norm")

# lnorm means lognormal
qqp(LenData$oocLen, "lnorm")

# gamma must be positive numbers
LenData$oocLen.l <- LenData$oocLen + 1
gamma <- fitdistr(LenData$oocLen.l, "gamma")
qqp(LenData$oocLen, "gamma", shape = gamma$estimate[[1]], rate = gamma$estimate[[2]])
```

### find best model
approximately normal distribution, continuous variable
```{r}
# null
oocNull <- glmer(oocLen ~ 1 + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
# full
oocFull <- glmer(oocLen ~ SocTreat + WingAvg + JHIII + RelTreat + DaysTreat + 
                  SocTreat*DaysTreat*RelTreat +
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
# 1 fixed effect, no interactions
ooc1 <- glmer(oocLen ~ SocTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc2 <- glmer(oocLen ~ WingAvg + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc3 <- glmer(oocLen ~ JHIII + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc4 <- glmer(oocLen ~ RelTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc5 <- glmer(oocLen ~ DaysTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
# 2 fixed effects, no interactions
ooc6 <- glmer(oocLen ~ SocTreat + WingAvg + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc7 <- glmer(oocLen ~ SocTreat + JHIII + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc8 <- glmer(oocLen ~ SocTreat + RelTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc9 <- glmer(oocLen ~ SocTreat + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc10 <- glmer(oocLen ~ WingAvg + JHIII + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc11 <- glmer(oocLen ~ WingAvg + RelTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc12 <- glmer(oocLen ~ WingAvg + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc13 <- glmer(oocLen ~ JHIII + RelTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc14 <- glmer(oocLen ~ JHIII + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc15 <- glmer(oocLen ~ RelTreat + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
# 3 fixed effects, no interactions
ooc16 <- glmer(oocLen ~ SocTreat + WingAvg + JHIII + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc17 <- glmer(oocLen ~ SocTreat + WingAvg + RelTreat +
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc18 <- glmer(oocLen ~ SocTreat + WingAvg + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc19 <- glmer(oocLen ~ SocTreat + JHIII + RelTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc20 <- glmer(oocLen ~ SocTreat + JHIII + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc21 <- glmer(oocLen ~ SocTreat + RelTreat + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc22 <- glmer(oocLen ~ WingAvg + JHIII + RelTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc23 <- glmer(oocLen ~ WingAvg + JHIII + DaysTreat +
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc24 <- glmer(oocLen ~ WingAvg + RelTreat + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc25 <- glmer(oocLen ~ JHIII + RelTreat + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
# 4 fixed effects, no interactions
ooc26 <- glmer(oocLen ~ SocTreat + WingAvg + JHIII + RelTreat +
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc27 <- glmer(oocLen ~ SocTreat + WingAvg + JHIII + DaysTreat +
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc28 <- glmer(oocLen ~ SocTreat + WingAvg + RelTreat + DaysTreat +
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc29 <- glmer(oocLen ~ SocTreat + JHIII + RelTreat + DaysTreat +
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc30 <- glmer(oocLen ~ WingAvg + JHIII + RelTreat + DaysTreat +
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
# 5 fixed effects, no interactions
ooc31 <- glmer(oocLen ~ SocTreat + JHIII + WingAvg + RelTreat + DaysTreat +
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
# 1 2-way interaction
ooc32 <- glmer(oocLen ~ SocTreat*WingAvg + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc33 <- glmer(oocLen ~ SocTreat*JHIII + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc34 <- glmer(oocLen ~ SocTreat*RelTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
#
ooc35 <- glmer(oocLen ~ SocTreat*DaysTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc36 <- glmer(oocLen ~ WingAvg*JHIII + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc37 <- glmer(oocLen ~ WingAvg*RelTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc38 <- glmer(oocLen ~ WingAvg*DaysTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc39 <- glmer(oocLen ~ JHIII*RelTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc40 <- glmer(oocLen ~ JHIII*DaysTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc41 <- glmer(oocLen ~ RelTreat*DaysTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
# 1 3-way interaction
ooc41 <- glmer(oocLen ~ SocTreat*RelTreat*DaysTreat + 
               (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
# all but one term
ooc42 <- glmer(oocLen ~ SocTreat + WingAvg + JHIII + RelTreat + DaysTreat + 
                SocTreat*WingAvg + SocTreat*DaysTreat + SocTreat*RelTreat + SocTreat*JHIII +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc43 <- glmer(oocLen ~ SocTreat + WingAvg + JHIII + RelTreat + DaysTreat +
                SocTreat*WingAvg + SocTreat*DaysTreat + SocTreat*RelTreat + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc44 <- glmer(oocLen ~ SocTreat + WingAvg + JHIII + RelTreat + DaysTreat +
                SocTreat*WingAvg + SocTreat*DaysTreat + SocTreat*JHIII + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc45 <- glmer(oocLen ~ SocTreat + WingAvg + JHIII + RelTreat + DaysTreat +
                SocTreat*WingAvg + SocTreat*RelTreat + SocTreat*JHIII + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc46 <- glmer(oocLen ~ SocTreat + WingAvg + JHIII + RelTreat + DaysTreat +
                SocTreat*DaysTreat + SocTreat*RelTreat + SocTreat*JHIII + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc47 <- glmer(oocLen ~ SocTreat + WingAvg + JHIII + RelTreat +
                SocTreat*WingAvg + SocTreat*RelTreat + SocTreat*JHIII +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc48 <- glmer(oocLen ~ SocTreat + WingAvg + JHIII + DaysTreat +
                SocTreat*WingAvg + SocTreat*DaysTreat + SocTreat*JHIII +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc49 <- glmer(oocLen ~ SocTreat + WingAvg + RelTreat + DaysTreat +
                SocTreat*WingAvg + SocTreat*DaysTreat + SocTreat*RelTreat + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc50 <- glmer(oocLen ~ SocTreat + JHIII + RelTreat + DaysTreat +
                SocTreat*DaysTreat + SocTreat*RelTreat + SocTreat*JHIII + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc51 <- glmer(oocLen ~ WingAvg + JHIII + RelTreat + DaysTreat +
                RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
# biological predictions
ooc52 <- glmer(oocLen ~ SocTreat + WingAvg + JHIII + RelTreat + DaysTreat +
                SocTreat*RelTreat + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc53 <- glmer(oocLen ~ SocTreat + WingAvg + RelTreat + DaysTreat +
                SocTreat*RelTreat + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc54 <- glmer(oocLen ~ SocTreat + WingAvg + JHIII + RelTreat + DaysTreat +
                SocTreat*RelTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc55 <- glmer(oocLen ~ SocTreat + WingAvg + RelTreat + DaysTreat +
                SocTreat*RelTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc56 <- glmer(oocLen ~ SocTreat + RelTreat + DaysTreat +
                RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc57 <- glmer(oocLen ~ SocTreat + WingAvg + JHIII +
                SocTreat*RelTreat + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc58 <- glmer(oocLen ~ SocTreat + WingAvg + JHIII + RelTreat + DaysTreat +
                SocTreat*RelTreat + RelTreat*DaysTreat + SocTreat*JHIII +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc59 <- glmer(oocLen ~ SocTreat + JHIII + RelTreat + DaysTreat +
                SocTreat*RelTreat + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc60 <- glmer(oocLen ~ SocTreat + JHIII + RelTreat + DaysTreat +
                SocTreat*RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
# full model -1 term
ooc61 <- glmer(oocLen ~ WingAvg + JHIII + RelTreat + DaysTreat +
                DaysTreat*RelTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc62 <- glmer(oocLen ~ SocTreat + JHIII + RelTreat + DaysTreat +
                SocTreat*DaysTreat*RelTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc63 <- glmer(oocLen ~ SocTreat + WingAvg + RelTreat + DaysTreat +
                SocTreat*DaysTreat*RelTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc64 <- glmer(oocLen ~ SocTreat + WingAvg + JHIII + DaysTreat +
                SocTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc65 <- glmer(oocLen ~ SocTreat + WingAvg + JHIII + RelTreat + 
                SocTreat*RelTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc66 <- glmer(oocLen ~ SocTreat + JHIII + DaysTreat + SocTreat*RelTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc67 <- glmer(oocLen ~ SocTreat + JHIII + SocTreat*RelTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc68 <- glmer(oocLen ~ SocTreat*JHIII + SocTreat*RelTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc69 <- glmer(oocLen ~ SocTreat*JHIII + SocTreat*RelTreat + DaysTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)
ooc70 <- glmer(oocLen ~ SocTreat + RelTreat + DaysTreat + JHIII +
                RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = gaussian(link = "identity"), data = LenData)

# best fit model = lowest AIC value
# WingAvg has 1 NA value, explaining error "models are not all fitted to the same data"
modSel <- model.sel(oocNull, oocFull, ooc1, ooc2, ooc3, ooc4, ooc5, ooc6, ooc7, ooc8, ooc9, ooc10, ooc11, ooc12, ooc13, ooc14, ooc15, ooc16, ooc17, ooc18, ooc19, ooc20, ooc21, ooc22, ooc23, ooc24, ooc25, ooc26, ooc27, ooc28, ooc29, ooc30, ooc31, ooc32, ooc33, ooc34, ooc35, ooc36, ooc37, ooc38, ooc39, ooc40, ooc41, ooc42, ooc43, ooc44, ooc45, ooc46, ooc47, ooc48, ooc49, ooc50, ooc51, ooc52, ooc23, ooc54, ooc55, ooc56, ooc57, ooc58, ooc60, ooc61, ooc62, ooc63, ooc64, ooc65, ooc66, ooc67, ooc68, ooc69, ooc70)

modSel
```

### summarize the model, overall p values
```{r}
# identify best model that does not produce an error message
mymodel <- ooc56

# summary for best model
summary(mymodel)
anova(mymodel)

# view estimates, confidence intervals, and p-values for model fixed and random effects
tab_model(mymodel)

# plot effects details to visualize results
plot(allEffects(mymodel))
```

### considerations
```{r}
# plot residuals
modelResiduals <- resid(mymodel, type = "pearson") # Extract standardized residuals
hist(modelResiduals)
qqp(modelResiduals, "norm")

# check for singularity
isSingular(mymodel)

# check for collinearity - value over 4ish signifies multicollinear, problematic, and potentially redundent
vif(mymodel)
```

### pairwise comparisons & p values
```{r}
# pairwise social vs. solitary at each collection day
# multiply p-values by 4 to get Bonferonni-corrected values

# early day 4
e <- LenData[LenData$RelTreat == "gas",]
e4 <- e[e$DaysTreat == "4",]
t.test(e4$oocLen[e4$SocTreat=="social"], e4$oocLen[e4$SocTreat=="solitary"], data = e4)

# early day 7
e7 <- e[e$DaysTreat == "7",]
t.test(e7$oocLen[e7$SocTreat=="social"], e7$oocLen[e7$SocTreat=="solitary"], data = e7)

# late day 4
l <- LenData[LenData$RelTreat == "offspring",]
l4 <- l[l$DaysTreat == "4",]
t.test(l4$oocLen[l4$SocTreat=="social"], l4$oocLen[l4$SocTreat=="solitary"], data = l4)

# late day 7
l7 <- l[l$DaysTreat == "7",]
t.test(l7$oocLen[l7$SocTreat=="social"], l7$oocLen[l7$SocTreat=="solitary"], data = l7)

```

### compare variances
```{r}
# early day 4
early4 <- LenData[LenData$RelTreat == "gas",]
early4 <- early4[early4$DaysTreat == "4",]
leveneTest(oocLen ~ SocTreat, data = early4)

# early day 7
early7 <- LenData[LenData$RelTreat == "gas",]
early7 <- early7[early7$DaysTreat == "7",]
leveneTest(oocLen ~ SocTreat, data = early7)

# late day 4
late4 <- LenData[LenData$RelTreat == "gas",]
late4 <- late4[late4$DaysTreat == "4",]
leveneTest(oocLen ~ SocTreat, data = late4)

# late day 7
late7 <- LenData[LenData$RelTreat == "gas",]
late7 <- late7[late7$DaysTreat == "4",]
leveneTest(oocLen ~ SocTreat, data = late7)
```





########################################################################################################







## oocyte resorption

### restructure dataframe to use in model
```{r}
# gather all ovariole stages into a single column
resData <- ovaryData %>%
  gather(ovR11, ovR12, ovR13, ovR14, ovR21, ovR22, ovR23, ovR24,
         key = "oocR",
         value = "resorb")
# remove na values
resData <- resData[!is.na(resData$resorb),]
# make integer values
resData$resorb[resData$resorb == "R"] <- 1
resData$resorb[resData$resorb == "U"] <- 0
resData$resorb <- as.factor(resData$resorb)
```

### decide on a distribution
binary response therefore binomial distribution
```{r}
# negative binomial must be discrete positive numbers
nbinom <- fitdistr(resData$resorb, "Negative Binomial")
qqp(resData$resorb, "nbinom", size = nbinom$estimate[[1]], mu = nbinom$estimate[[2]])

# poisson must be positive whole numbers
poisson <- fitdistr(resData$resorb, "Poisson")
qqp(resData$resorb, "pois", poisson$estimate)
```

### add oocAvg to use as covariate
```{r, echo = FALSE}
# average the oocyte lengths for each bee 
for (i in 1:nrow(resData)) {
  resData$oocAvg[i] <- mean(c(resData$Ooc11[i],
                              resData$Ooc12[i],
                              resData$Ooc13[i],
                              resData$Ooc14[i],
                              resData$Ooc21[i],
                              resData$Ooc22[i],
                              resData$Ooc23[i],
                              resData$Ooc24[i]),
                              na.rm = TRUE) }
```

## find best model
binomial distribution, binary response
```{r}
# null
resNull <- glmer(resorb ~ 1 + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
# full
resFull <- glmer(resorb ~ SocTreat + WingAvg + JHIII + RelTreat + DaysTreat + oocAvg +
                  SocTreat*DaysTreat*RelTreat +
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
# 1 fixed effect, no interactions
res1 <- glmer(resorb ~ SocTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res2 <- glmer(resorb ~ WingAvg + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res3 <- glmer(resorb ~ JHIII + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res4 <- glmer(resorb ~ RelTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res5 <- glmer(resorb ~ DaysTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res71 <-glmer(resorb ~ oocAvg + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
# 2 fixed effects, no interactions
res6 <- glmer(resorb ~ SocTreat + WingAvg + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res7 <- glmer(resorb ~ SocTreat + JHIII + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res8 <- glmer(resorb ~ SocTreat + RelTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res9 <- glmer(resorb ~ SocTreat + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res10 <- glmer(resorb ~ WingAvg + JHIII + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res11 <- glmer(resorb ~ WingAvg + RelTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res12 <- glmer(resorb ~ WingAvg + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res13 <- glmer(resorb ~ JHIII + RelTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res14 <- glmer(resorb ~ JHIII + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res15 <- glmer(resorb ~ RelTreat + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res72 <- glmer(resorb ~ oocAvg + SocTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res73 <- glmer(resorb ~ oocAvg + WingAvg + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res74 <- glmer(resorb ~ oocAvg + JHIII + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res75 <- glmer(resorb ~ oocAvg + RelTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res76 <- glmer(resorb ~ oocAvg + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
# 3 fixed effects, no interactions
res16 <- glmer(resorb ~ SocTreat + WingAvg + JHIII + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res17 <- glmer(resorb ~ SocTreat + WingAvg + RelTreat +
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res18 <- glmer(resorb ~ SocTreat + WingAvg + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res19 <- glmer(resorb ~ SocTreat + JHIII + RelTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res20 <- glmer(resorb ~ SocTreat + JHIII + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res21 <- glmer(resorb ~ SocTreat + RelTreat + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res22 <- glmer(resorb ~ WingAvg + JHIII + RelTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res23 <- glmer(resorb ~ WingAvg + JHIII + DaysTreat +
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res24 <- glmer(resorb ~ WingAvg + RelTreat + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res25 <- glmer(resorb ~ JHIII + RelTreat + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res77 <- glmer(resorb ~ oocAvg + RelTreat + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res78 <- glmer(resorb ~ oocAvg + SocTreat + RelTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res79 <- glmer(resorb ~ oocAvg + SocTreat + DaysTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res80 <- glmer(resorb ~ oocAvg + JHIII + SocTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res81 <- glmer(resorb ~ oocAvg + RelTreat + JHIII + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
# 4 fixed effects, no interactions
res26 <- glmer(resorb ~ SocTreat + WingAvg + JHIII + RelTreat +
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res27 <- glmer(resorb ~ SocTreat + WingAvg + JHIII + DaysTreat +
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res28 <- glmer(resorb ~ SocTreat + WingAvg + RelTreat + DaysTreat +
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res29 <- glmer(resorb ~ SocTreat + JHIII + RelTreat + DaysTreat +
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res30 <- glmer(resorb ~ WingAvg + JHIII + RelTreat + DaysTreat +
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
# 5 fixed effects, no interactions
res31 <- glmer(resorb ~ SocTreat + JHIII + WingAvg + RelTreat + DaysTreat +
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
# 1 2-way interaction
res32 <- glmer(resorb ~ SocTreat*WingAvg + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res33 <- glmer(resorb ~ SocTreat*JHIII + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res34 <- glmer(resorb ~ SocTreat*RelTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res35 <- glmer(resorb ~ SocTreat*DaysTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res36 <- glmer(resorb ~ WingAvg*JHIII + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res37 <- glmer(resorb ~ WingAvg*RelTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res38 <- glmer(resorb ~ WingAvg*DaysTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res39 <- glmer(resorb ~ JHIII*RelTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res40 <- glmer(resorb ~ JHIII*DaysTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res41 <- glmer(resorb ~ RelTreat*DaysTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
# 1 3-way interaction
res41 <- glmer(resorb ~ SocTreat*RelTreat*DaysTreat + 
               (1|NatCol) + (1|BeeID), family = binomial, data = resData)
# all but one term
res42 <- glmer(resorb ~ SocTreat + WingAvg + JHIII + RelTreat + DaysTreat + 
                SocTreat*WingAvg + SocTreat*DaysTreat + SocTreat*RelTreat + SocTreat*JHIII +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res43 <- glmer(resorb ~ SocTreat + WingAvg + JHIII + RelTreat + DaysTreat +
                SocTreat*WingAvg + SocTreat*DaysTreat + SocTreat*RelTreat + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res44 <- glmer(resorb ~ SocTreat + WingAvg + JHIII + RelTreat + DaysTreat +
                SocTreat*WingAvg + SocTreat*DaysTreat + SocTreat*JHIII + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res45 <- glmer(resorb ~ SocTreat + WingAvg + JHIII + RelTreat + DaysTreat +
                SocTreat*WingAvg + SocTreat*RelTreat + SocTreat*JHIII + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res46 <- glmer(resorb ~ SocTreat + WingAvg + JHIII + RelTreat + DaysTreat +
                SocTreat*DaysTreat + SocTreat*RelTreat + SocTreat*JHIII + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res47 <- glmer(resorb ~ SocTreat + WingAvg + JHIII + RelTreat +
                SocTreat*WingAvg + SocTreat*RelTreat + SocTreat*JHIII +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res48 <- glmer(resorb ~ SocTreat + WingAvg + JHIII + DaysTreat +
                SocTreat*WingAvg + SocTreat*DaysTreat + SocTreat*JHIII +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res49 <- glmer(resorb ~ SocTreat + WingAvg + RelTreat + DaysTreat +
                SocTreat*WingAvg + SocTreat*DaysTreat + SocTreat*RelTreat + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res50 <- glmer(resorb ~ SocTreat + JHIII + RelTreat + DaysTreat +
                SocTreat*DaysTreat + SocTreat*RelTreat + SocTreat*JHIII + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res51 <- glmer(resorb ~ WingAvg + JHIII + RelTreat + DaysTreat +
                RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
# biological predictions
res52 <- glmer(resorb ~ SocTreat + WingAvg + JHIII + RelTreat + DaysTreat +
                SocTreat*RelTreat + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res53 <- glmer(resorb ~ SocTreat + WingAvg + RelTreat + DaysTreat +
                SocTreat*RelTreat + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res54 <- glmer(resorb ~ SocTreat + WingAvg + JHIII + RelTreat + DaysTreat +
                SocTreat*RelTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res55 <- glmer(resorb ~ SocTreat + WingAvg + RelTreat + DaysTreat +
                SocTreat*RelTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res56 <- glmer(resorb ~ SocTreat + RelTreat + DaysTreat +
                RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res57 <- glmer(resorb ~ SocTreat + WingAvg + JHIII +
                SocTreat*RelTreat + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res58 <- glmer(resorb ~ SocTreat + WingAvg + JHIII + RelTreat + DaysTreat +
                SocTreat*RelTreat + RelTreat*DaysTreat + SocTreat*JHIII +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res59 <- glmer(resorb ~ SocTreat + JHIII + RelTreat + DaysTreat +
                SocTreat*RelTreat + RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res60 <- glmer(resorb ~ SocTreat + JHIII + RelTreat + DaysTreat +
                SocTreat*RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
# full model -1 term
res61 <- glmer(resorb ~ WingAvg + JHIII + RelTreat + DaysTreat +
                DaysTreat*RelTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res62 <- glmer(resorb ~ SocTreat + JHIII + RelTreat + DaysTreat +
                SocTreat*DaysTreat*RelTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res63 <- glmer(resorb ~ SocTreat + WingAvg + RelTreat + DaysTreat +
                SocTreat*DaysTreat*RelTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res64 <- glmer(resorb ~ SocTreat + WingAvg + JHIII + DaysTreat +
                SocTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res65 <- glmer(resorb ~ SocTreat + WingAvg + JHIII + RelTreat + 
                SocTreat*RelTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res66 <- glmer(resorb ~ SocTreat + JHIII + DaysTreat + SocTreat*RelTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res67 <- glmer(resorb ~ SocTreat + JHIII + SocTreat*RelTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res68 <- glmer(resorb ~ SocTreat*JHIII + SocTreat*RelTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res69 <- glmer(resorb ~ SocTreat*JHIII + SocTreat*RelTreat + DaysTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res70 <- glmer(resorb ~ SocTreat + RelTreat + DaysTreat + JHIII +
                RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)
res82 <- glmer(resorb ~ SocTreat + RelTreat + DaysTreat + JHIII + oocAvg +
                RelTreat*DaysTreat +
                (1|NatCol) + (1|BeeID), family = binomial, data = resData)

# best fit model = lowest AIC value
# WingAvg has 1 NA value, explaining error "models are not all fitted to the same data"
modSel <- model.sel(resNull, resFull, res1, res2, res3, res4, res5, res6, res7, res8, res9, res10, res11, res12, res13, res14, res15, res16, res17, res18, res19, res20, res21, res22, res23, res24, res25, res26, res27, res28, res29, res30, res31, res32, res33, res34, res35, res36, res37, res38, res39, res40, res41, res42, res43, res44, res45, res46, res47, res48, res49, res50, res51, res52, res23, res54, res55, res56, res57, res58, res60, res61, res62, res63, res64, res65, res66, res67, res68, res69, res70, res71, res72, res73, res74, res75, res76, res77, res78, res79, res80, res81, res82)

modSel
```

### summarize the model, overall p values
```{r}
# identify best model that does not produce an error message
mymodel <- res34

# summary for best model
summary(mymodel)
anova(mymodel)

# view estimates, confidence intervals, and p-values for model fixed and random effects
tab_model(mymodel)

# plot effects details to visualize results
plot(allEffects(mymodel))
```

### considerations
```{r}
# plot residuals
modelResiduals <- resid(mymodel, type = "pearson") # Extract standardized residuals
hist(modelResiduals)
qqp(modelResiduals, "norm")

# check for singularity
isSingular(mymodel)

# check for collinearity - value over 4ish signifies multicollinear, problematic, and potentially redundent
vif(mymodel)

# check for overdispersion - p < 0.05 = oversdispersed and invalid model
overdisp_fun <- function(mymodel) {
    rdf <- df.residual(mymodel)
    rp <- residuals(mymodel,type="pearson")
    Pearson.chisq <- sum(rp^2)
    prat <- Pearson.chisq/rdf
    pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
    c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}
overdisp_fun(mymodel)
```

### pairwise comparisons & p-values
```{r}
# posthoc tukey's test comparing of social-solitary queens within each collection stage
posthoc1 <- lsmeans(mymodel, pairwise ~ SocTreat|RelTreat, adjust = "tukey")
summary(posthoc1)
```
