---
title: "btRepro"
output: pdf_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages
```{r, include = FALSE}
library(tidyverse)
library(ggplot2)
library(plyr)
library(lme4)
library(car)
library(MuMIn)
```

# import data from cleanData file
```{r}
# assumes working directory is where the original datasheet lives
# getwd()
btData <- read.csv("./clean2.csv", header = TRUE)
btData

# make social treatment a factor with ordered levels
btData$socTreat <- factor(btData$socTreat, levels = c("Solitary", "Social")) 
```

# visualize days to eggs
```{r}
# subset to only data i want to graph
plotData <- btData[!is.na(btData$daysToEggs),]

# count sample size
sampleSizes <- plotData %>% group_by(history, socTreat) %>% tally()
sampleSizes

# summarize the data and calculate mean, sd, sem
summ <- ddply(plotData, 
                c("history", "nest", "socTreat"), 
                summarise,
                mean = mean(daysToEggs), 
                sd = sd(daysToEggs),
                sem = sd(daysToEggs)/sqrt(length(daysToEggs)))

# split summ into lines
solsol <- summ[c(1, 5),]
solsol$history <- as.character(solsol$history)
solsol <- solsol[order(solsol$history),]
solsol$history <- as.factor(solsol$history)
solsoc <- summ[c(1, 6),]
solsoc$history <- as.character(solsoc$history)
solsoc <- solsoc[order(solsoc$history),]
solsoc$history <- as.factor(solsoc$history)
socsol <- summ[c(2, 3),]
socsol$history <- as.character(socsol$history)
socsol <- socsol[order(socsol$history),]
socsol$history <- as.factor(socsol$history)
socsoc <- summ[c(2, 4),]
socsoc$history <- as.character(socsoc$history)
socsoc <- socsoc[order(socsoc$history),]
socsoc$history <- as.factor(socsoc$history)

# line graph
ggplot() +
  geom_line(data = solsol, aes(x = nest, y = mean, group = 1), linetype = "solid") +
  geom_line(data = solsoc, aes(x = nest, y = mean, group = 1), linetype = "solid") +
  geom_line(data = socsol, aes(x = nest, y = mean, group = 1), linetype = "dotted") +
  geom_line(data = socsoc, aes(x = nest, y = mean, group = 1), linetype = "dotted") +
  geom_errorbar(data = summ, aes(x = nest, y = mean, fill = socTreat, ymin = mean - sem, ymax = mean + sem),
                width = 0.05) +
  geom_point(data = summ, aes(x = nest, y = mean, shape = socTreat), size = 5) +
  scale_color_grey(start = 0, end = .7) +
  scale_shape_manual(values = c(16, 21)) +
  labs(x = "nest", y = "days until first eggs", 
      color = "social treatment", shape = "social status") +
  theme(panel.background = element_blank(), axis.line = element_line(colour = "black"), 
        text = element_text(size=15))
```
# mixed models

```{r}
library(sjPlot)
library(effects)
library(lsmeans)
```

## decide on a distribution
### my data is not normal, lnorm fits best
```{r}
dayData <- btData[!is.na(btData$daysToEggs),]

# normal distribution
qqp(dayData$daysToEggs, "norm")

# lnorm means lognormal
qqp(dayData$daysToEggs, "lnorm")

# negative binomial must be discrete positive numbers
nbinom <- fitdistr(dayData$daysToEggs, "Negative Binomial")
qqp(dayData$daysToEggs, "nbinom", size = nbinom$estimate[[1]], mu = nbinom$estimate[[2]])

# poisson must be positive whole numbers
poisson <- fitdistr(dayData$daysToEggs, "Poisson")
qqp(dayData$daysToEggs, "pois", lambda = poisson$estimate)

# gamma must be positive numbers
gamma <- fitdistr(dayData$daysToEggs, "gamma")
qqp(dayData$daysToEggs, "gamma", shape = gamma$estimate[[1]], rate = gamma$estimate[[2]])

###################################################################################

# plot the social treatments separately to verify they still fit lnorm

# subset the data into social and solitary
early <- dayData[dayData$RelTreat == "gas",]
late <- dayData[dayData$RelTreat == "offspring",]

# poisson must be positive whole numbers
poisson <- fitdistr(late$daysToEggs, "Poisson")
qqp(late$daysToEggs, "pois", lambda = poisson$estimate)

# transform the data
dayData$daysToEggs.l <- log(dayData$daysToEggs + 0.01)

```

### find best model
lnorm distribution, discrete variable
```{r}
# models
daysNull <- glmer(daysToEggs.l ~ 1 + 
               (1|qNatCol), family = gaussian(link = "log"), data = dayData)
daysFull <- glmer(daysToEggs.l ~ nest * socTreat * history + 
               (1|qNatCol), family = gaussian(link = "log"), data = dayData)
# all but 1
days1 <- glmer(daysToEggs.l ~ nest * socTreat + 
               (1|qNatCol), family = gaussian(link = "log"), data = dayData)
days2 <- glmer(daysToEggs.l ~ nest * history + 
               (1|qNatCol), family = gaussian(link = "log"), data = dayData)
days3 <- glmer(daysToEggs.l ~ socTreat * history + 
               (1|qNatCol), family = gaussian(link = "log"), data = dayData)
# only 1
days4 <- glmer(daysToEggs.l ~ nest + 
               (1|qNatCol), family = gaussian(link = "log"), data = dayData)
days5 <- glmer(daysToEggs.l ~ socTreat + 
               (1|qNatCol), family = gaussian(link = "log"), data = dayData)
days6 <- glmer(daysToEggs.l ~ history + 
               (1|qNatCol), family = gaussian(link = "log"), data = dayData)
# all, no interactions
days7 <- glmer(daysToEggs.l ~ nest + socTreat + history + 
               (1|qNatCol), family = gaussian(link = "log"), data = dayData)
# 2, no interactions
days8 <- glmer(daysToEggs.l ~ nest + socTreat + 
               (1|qNatCol), family = gaussian(link = "log"), data = dayData)
days9 <- glmer(daysToEggs.l ~ nest + history + 
               (1|qNatCol), family = gaussian(link = "log"), data = dayData)
days10 <- glmer(daysToEggs.l ~ socTreat + history + 
               (1|qNatCol), family = gaussian(link = "log"), data = dayData)

modSel <- model.sel(daysNull, daysFull, days1, days2, days3, days4, days5, days6, days7, days8, days9, days10)
modSel
```

### summarize the model, overall p values
```{r}
# identify best model that does not produce an error message
mymodel <- days1

# summary for best model
summary(mymodel)
anova(mymodel)

# view estimates, confidence intervals, and p-values for model fixed and random effects
tab_model(mymodel)

# plot effects details to visualize results
plot(allEffects(mymodel))
```

### considerations
```{r}
# plot residuals
modelResiduals <- resid(mymodel, type = "pearson") # Extract standardized residuals
hist(modelResiduals)
qqp(modelResiduals, "norm")

# check for singularity
isSingular(mymodel)

# check for collinearity - value over 4ish signifies multicollinear, problematic, and potentially redundent
vif(mymodel)
```

### pairwise comparisons & p-values
```{r}
# posthoc tukey's test comparing of social-solitary queens within each collection stage
posthoc1 <- lsmeans(mymodel, pairwise ~ nest|socTreat, adjust = "tukey")
summary(posthoc1)

# compare variances

# levene's test for equal variances
# first nest
first <- dayData[dayData$nest == "First",]
leveneTest(daysToEggs ~ socTreat, data = first)
first

# second nest
second <- dayData[dayData$nest == "Second",]
leveneTest(daysToEggs ~ socTreat, data = second)
```

######
###
#
######
###
#
######
###
#
######
###
#
######
###
#
######
###
#
######
###
#
######
###
#
######
###
#
######
###
#
#######
###
#

# visualize number of brood
```{r}
# subset to only data i want to graph
plotData <- btData[!is.na(btData$nAllBrood),]

# count sample size
sampleSizes <- plotData %>% group_by(history, socTreat) %>% tally()
sampleSizes

# summarize the data and calculate mean, sd, sem
summ <- ddply(plotData, 
                c("history", "nest", "socTreat"), 
                summarise,
                mean = mean(nAllBrood), 
                sd = sd(nAllBrood),
                sem = sd(nAllBrood)/sqrt(length(nAllBrood)))
summ

# split summ into lines
solsol <- summ[c(1, 5),]
solsol$history <- as.character(solsol$history)
solsol <- solsol[order(solsol$history),]
solsol$history <- as.factor(solsol$history)
solsoc <- summ[c(1, 6),]
solsoc$history <- as.character(solsoc$history)
solsoc <- solsoc[order(solsoc$history),]
solsoc$history <- as.factor(solsoc$history)
socsol <- summ[c(2, 3),]
socsol$history <- as.character(socsol$history)
socsol <- socsol[order(socsol$history),]
socsol$history <- as.factor(socsol$history)
socsoc <- summ[c(2, 4),]
socsoc$history <- as.character(socsoc$history)
socsoc <- socsoc[order(socsoc$history),]
socsoc$history <- as.factor(socsoc$history)

# line graph
ggplot() +
  geom_line(data = solsol, aes(x = nest, y = mean, group = 1), linetype = "solid") +
  geom_line(data = solsoc, aes(x = nest, y = mean, group = 1), linetype = "solid") +
  geom_line(data = socsol, aes(x = nest, y = mean, group = 1), linetype = "dotted") +
  geom_line(data = socsoc, aes(x = nest, y = mean, group = 1), linetype = "dotted") +
  geom_errorbar(data = summ, aes(x = nest, y = mean, fill = socTreat, ymin = mean - sem, ymax = mean + sem),
                #position = position_dodge(.2),
                width = 0.05) +
  geom_point(data = summ, aes(x = nest, y = mean, shape = socTreat), size = 5) +
  scale_color_grey(start = 0, end = .7) +
    scale_shape_manual(values = c(16, 21)) +
  # scale_shape_manual(values = c(15, 16, 17),
  #                   labels = c("NA", "was previously social", "was previously solitary")) + 
  labs(x = "nest", y = "number of brood items", 
      color = "social treatment", shape = "social treatment") +
  # guides(col = guide_legend(override.aes = list(shape = 15, size = 6.5)),
  #        shape = guide_legend(override.aes = list(shape = c(0, 1, 2)))) +
  theme(panel.background = element_blank(), axis.line = element_line(colour = "black"), 
        text = element_text(size=15))
```

# mixed models
normally distributed
```{r}
# normal distribution
qqp(btData$nAllBrood, "norm")

# lnorm means lognormal
qqp(btData$nAllBrood, "lnorm")

# negative binomial must be discrete positive numbers
nbinom <- fitdistr(btData$nAllBrood, "Negative Binomial")
qqp(btData$nAllBrood, "nbinom", size = nbinom$estimate[[1]], mu = nbinom$estimate[[2]])

# poisson must be positive whole numbers
poisson <- fitdistr(btData$nAllBrood, "Poisson")
qqp(btData$nAllBrood, "pois", lambda = poisson$estimate)

# gamma must be positive numbers
gamma <- fitdistr(btData$nAllBrood, "gamma")
qqp(btData$nAllBrood, "gamma", shape = gamma$estimate[[1]], rate = gamma$estimate[[2]])
```

### find best model
normal distribution
```{r}
# models
bNull <- glmer(nAllBrood ~ 1 + 
               (1|qNatCol), family = gaussian(link = "identity"), data = btData)
bFull <- glmer(nAllBrood ~ nest * socTreat * history + 
               (1|qNatCol), family = gaussian(link = "identity"), data = btData)
# all but 1
b1 <- glmer(nAllBrood ~ nest * socTreat + 
               (1|qNatCol), family = gaussian(link = "identity"), data = btData)
b2 <- glmer(nAllBrood ~ nest * history + 
               (1|qNatCol), family = gaussian(link = "identity"), data = btData)
b3 <- glmer(nAllBrood ~ socTreat * history + 
               (1|qNatCol), family = gaussian(link = "identity"), data = btData)
# only 1
b4 <- glmer(nAllBrood ~ nest + 
               (1|qNatCol), family = gaussian(link = "identity"), data = btData)
b5 <- glmer(nAllBrood ~ socTreat + 
               (1|qNatCol), family = gaussian(link = "identity"), data = btData)
b6 <- glmer(nAllBrood ~ history + 
               (1|qNatCol), family = gaussian(link = "identity"), data = btData)
# all, no interactions
b7 <- glmer(nAllBrood ~ nest + socTreat + history + 
               (1|qNatCol), family = gaussian(link = "identity"), data = btData)
# 2, no interactions
b8 <- glmer(nAllBrood ~ nest + socTreat + 
               (1|qNatCol), family = gaussian(link = "identity"), data = btData)
b9 <- glmer(nAllBrood ~ nest + history + 
               (1|qNatCol), family = gaussian(link = "identity"), data = btData)
b10 <- glmer(nAllBrood ~ socTreat + history + 
               (1|qNatCol), family = gaussian(link = "identity"), data = btData)

modSel <- model.sel(bNull, bFull, b1, b2, b3, b4, b5, b6, b7, b8, b9, b10)
modSel
```

### summarize the model, overall p-values
```{r}
# identify best model that does not produce an error message
mymodel <- b3

# summary for best model
summary(mymodel)
anova(mymodel)

# view estimates, confidence intervals, and p-values for model fixed and random effects
tab_model(mymodel)

# plot effects details to visualize results
plot(allEffects(mymodel))
```

### considerations
```{r}
# plot residuals
modelResiduals <- resid(mymodel, type = "pearson") # Extract standardized residuals
hist(modelResiduals)
qqp(modelResiduals, "norm")

# check for singularity
isSingular(mymodel)

# check for collinearity - value over 4ish signifies multicollinear, problematic, and potentially redundent
vif(mymodel)
```

### pairwise comparisons & p-values
```{r}
# posthoc tukey's tests comparing solitary and social groups within each social history
posthoc1 <- lsmeans(mymodel, pairwise ~ socTreat|history, adjust = "tukey")
summary(posthoc1)
```

